Measuring contrast processing in the visual system using SSVEP

### 2024-05-09 review.tex
Code to make R+H plots -

The complexity of even a simple simulation of the frequency-domain
signal arising from non-linear interactions is intriguing. Presumably,
the signal measured from early visual cortex is the result of a cascade
of nonlinear retinal and cortical operations up to that point. It
therefore contains a `signature' of the computational nature, order and
parameters of those operations - including the shape of the transducer
functions and the computations involved in signal combination. In
principle, that information could be recovered from the SSVEP signal - a
possibility recognised in the early days of the technique (Regan \&
Regan, 1988). Although characterising the complete set of computations
along entire processing pathway is challenging, careful parametric
variation of the input stimuli does allow us to fit models of early
visual gain control using SSVEP data. For example, Tsai \emph{et al.}
(2012) demonstrated that a gain control model gave a good account of the
pattern of intermodulation responses produced by two overlaid patterns
flickering at different frequencies. This was achieved by passing full
stimulus waveforms through the transducer nonlinearity, and calculating
the Fourier spectrum of the model output. Our own work on signal
combination across eyes and space similarly demonstrated close
correspondence between the predictions of a computational model and
empirical data in humans (Baker \& Wade, 2017). More detailed modelling
of intracortical recordings (Groen \emph{et al.}, 2022) has revealed
details of the timecourse of gain control effects, specifically that
normalization is delayed slightly relative to the initial visual
response.

@@

Because SSVEP responses at detection threshold are very small,
estimating a threshold is achieved by measuring the contrast response
function at relatively high levels, and extrapolating back along the
function (either contrast vs response measured at a constant spatial
frequency or spatial frequency vs response at a constant contrast level)
to estimate its intercept with the x-axis (see
Figure~\ref{fig-sweepvep}). This contrast level was shown to correspond
approximately with psychophysically measured detection thresholds
(Norcia \emph{et al.}, 1986).

@@

\section{Attention}\label{attention}

Attention exerts a profound influence on visual performance: for
example, allowing people to attend covertly to a spatial location
improves their performance on a target detection task significantly
(Bashinski \& Bacharach, 1980; Posner, 1980; Carrasco \emph{et al.},
2000; Cameron \emph{et al.}, 2002; Morrone \emph{et al.}, 2004; Pestilli
\emph{et al.}, 2009). In principle, this enhancement may be driven both
by modulation of the underlying signal or noise characteristics, or by
additional decision-theoretic factors such as reduction in spatial
uncertainty (Petrov \emph{et al.}, 2006; Gould \emph{et al.}, 2007).
Early experiments with non-human primates showed little evidence for
attentionally-driven changes in neuronal spike rates (Luck \emph{et
al.}, 1997; McAdams \& Maunsell, 1999; Mehta \emph{et al.},
2000\emph{a}, 2000\emph{b}; Marcus \& Van Essen, 2002), but with the
advent of spatially-resolved human brain imaging methods in the late
1990s it became apparent that spatial attention was linked to frank
changes in both fMRI BOLD responses (Tootell \emph{et al.}, 1998; Somers
\emph{et al.}, 1999; Gandhi \emph{et al.}, 1999; Brefczynski \& DeYoe,
1999; Kastner \emph{et al.}, 1999; Buracas \& Boynton, 2007; Silver
\emph{et al.}, 2007; Li \emph{et al.}, 2008; Murray, 2008) and EEG
signals (\textbf{MullerHillyard2000?}).

Electrophysiological and psychophysical measurements of the effect of
attention on both luminance and chromatic contrast have strongly
implicated gain control as an underlying mechanism (Lu \& Dosher, 1998;
Di Russo \& Spinelli, 1999\emph{a}; Di Russo \emph{et al.},
2001\emph{b}). These effects also appear to differ between chromatic and
achromatic pathways (Di Russo \& Spinelli, 1999\emph{b}) - perhaps as a
result of the different levels of nonlinear gain control in the early
pre-cortical magno-, parvo- and konio-cellular pathways (Derrington \&
Lennie, 1984; Kaplan \& Shapley, 1986; Lee \emph{et al.}, 1990; Solomon
\& Lennie, 2005). In the late 2000s a comprehensive theoretical model
for attentional modulation was developed that framed it as a gain
control computation (Reynolds \& Heeger, 2009; Boynton, 2009). This
framework has proven to be influential - explaining a wide range of
phenomena from the earlier literature and demonstrating subtle
interactions between the size of the attentional `spotlight' and the
stimulus configuration which rationalise many apparent contradictions in
the literature. Direct measurements of attentional modulation of
achromatic SSVEP signals are broadly consistent with this model
Martinovic \& Andersen (2018) and confirm the relatively weaker role of
spatial attention on responses driven by chromatic stimuli -
particularly those that isolate the opponent S-(L+M) cone pathway
Highsmith \& Crognale (2010). The gain control model of attention can
also be extended to SSVEP studies of feature-based attention which show
that the modulatory effects can be targeted to the most informative
neuronal populations (Verghese \emph{et al.}, 2012).

@@

One widespread recent application of the SSVEP technique is in the
design of brain-computer interfaces, which seek to control some aspect
of a computer using neural signals. The high SNR and frequency
resolution of SSVEPs make them an ideal candidate for this approach.
Typical studies may involve presenting an array of stimuli at different
contrast flicker frequencies, and having the participant select one
either by overt attention (i.e.~shifting fixation to foveate the
selected stimulus) or covert attention (i.e.~deploying attention to one
stimulus whilst keeping fixated) (\textbf{Middendorf?}). BEcause EEG
signals are highly sensitive to both visual field position (Ales
\emph{et al.}, 2010) and attentional state (Lauritzen \emph{et al.},
2010; Verghese \emph{et al.}, 2012), the response to the selected
stimulus will typically increase relative to the others, allowing it to
be identified by an on-line algorithm. Because more than one stimulus
frequency is typically present, this modulation and the associated
changes in gain control, will affect the entire pattern of self- and
intermodulation terms.

This approach is primarily useful in situations that require the BCI to
distinguish from among a small set of possibilties : For example, in
early work, visual stimuli representing `Left' and `Right' commands in a
flight simulator were distinguished robustly (\textbf{Middendorf?}).
Although SSVEP-based BCI interfaces typically do use contrast flicker,
work in this field has largely focused on optimising the stimuli or
decoders to increase decoding performance (e.g.~\ldots{} ) and reducing
visual fatigue associated with the long-term presentation of arrays of
high-contrast flicker (e.g.~\ldots..) rather than studying visual
contrast processing \textit{per se}.

\section{Conclusions}\label{conclusions}

\emph{2: Sweeps and CRFs. Measurements of contrast sensitivity,
extrapolating the sweep to zero response to get t'hold. Infants and
adults?}

\emph{3: Measurements of modulation. Figures adapted from other papers:
Attention to space, attention to features, adaptation (?).
masking/surround suppression\textbackslash{}}

\emph{4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)}

\emph{5: Future directions (decoding in frequency domain?, animals?
BCI?)}

\phantomsection\label{anchor-2}{}History

\phantomsection\label{anchor-3}{}1. The basics of SSVEP and contrast
sensitivity including a history of both fields

SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
response evoked in the brain by visual stimuli flickering at a constant
frequency (Regan, 1966).

Contrast Sensitivity: The ability to detect differences in luminance
between an object and its background (Campbell \& Green, 1965).

Regan, D. (1966). Some characteristics of average steady-state and
transient responses evoked by modulated light. Electroencephalography
and clinical neurophysiology, 20(3), 238-248.

Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
affecting visual resolution. J Physiology, 181, 576-593.

Check Regan paper for earlier (e.g.~EEG refs). Norcia review will be
helpful!

(From Tyler / Levi / Apkarian paper):

6. Regan, D.: Rapid methods for refracting the eye and assessing the
visual acuity in amblyopia using steady-state visual evoked potentials.
In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
Developments, Oxford, 1977, Clarendon Press, pp.~418-426.

7. Fricker, S.J.: Narrow-band filter techniques for the detection and
measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
14:411, 1962.

8. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
W.S., and Verings, F.T.H.: Objective determination of response to
modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.

9. Regan, D.: Latencies of evoked potentials to flicker and to pattern
speedily estimated by simultaneous stimulation method,
Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.

10. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
frequency tuning of electrical responses from the human.\ldots{}

{[}Around here perhaps a section about what gain control is, mentioning
other methods as well including psychophysics, MRI, electrophysiology,
and other EEG markers including ERPs and evoked gamma band oscillations.
Maybe outline the Heeger gain control model and its cousins.{]}

{[}Sure. But in fact gain control is only a small part of this story -
especially in the early days. They were, I think, more interested in
using SSVEP to measure absolute contrast sensitivity and to get lower
bounds on things like infant visual development. Gain control might be
better as a separate section later.{]}

\phantomsection\label{anchor-4}{}2. Spatial, temporal frequency and
contrast sensitivity measurements

Basically using SSVEP to measure a cortical output amplitude for any
given input contrast. You can vary parameters like SF, TF, position,
color and of course contrast. Early on people realised that you can
`sweep' the stimulus to get a CRF. You broadly get a line in log
contrast space if you do that (Tyler) - then you can extrapolate that
line down to zero response to estimate the threshold. That doesn't
\textgreater quite\textless{} work but it's pretty close.

You can also use this for `difficult' populations like babies. One
interesting story was about how SSVEP become a replacement for
preferential looking (which was the other way of looking at infant
visual development). See e.g.~Davida Teller. SSVEPs allowed people to
make objective measurements of contrast sensitivity development and
deduce that the visual system was more mature (e.g.~more functional) in
infancy than previously expected. Also measures of colour sensitivity.
Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
86,88,90 - mentioned in Regan's nice autobiography :

In parallel of course, people were using frequency tagging to do single
unit work - the 1F vs 2F simple/complex cell classification scheme was
all about this (Lennie and others).

\phantomsection\label{anchor-5}{}3. SSVEP in functional localization

Techniques like fMRI have been combined with SSVEP to achieve more
precise spatial localization (Di Russo et al., 2007). See however the
Ales cruciform paper.

Ales and Norcia:... (showed that people's intuition about V1 upper /
lower v.f reversing polarity wasn't really correct). This feels a little
outside our scope though... We have also combined SSVEP with source
imaging techniques to probe responses in different cortical locations
(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
Norcia figure/ground...) and then a host of later work from that lab and
others.

\phantomsection\label{anchor-6}{}4. SSVEP and contrast gain control
including adaptation, masking, and attention

{[}The story goes: Up to the lte 90s people were primarily interested in
measuring contrast sensitivity - the shape of the response function was
assumed to be basically log-linear - and they fit it with straight lines
to extrapolate back to zero response. But then (once Heeger's 1992 paper
had sunk in - see also stuff like Shapley and Victor 1981
\href{https://www.zotero.org/google-docs/?zAmtgj}{(Shapley \& Victor,
1981)}), people started thinking about gain control - Candy and Norcia
in about 1999, Porciatti, probably a load of Tyler papers that I don't
even know about...{]}. Probably look in p

And then people worked out that if you can use SSVEP to measure contrast
responses, you can also use it to measure things that modulate contrast
responses. These include adaptation, masking, suppression, attention
(feature and space), clinical things.

\phantomsection\label{anchor-7}{}Adaptation:

Continuous exposure to high-contrast patterns reduces contrast
sensitivity, which can be measured using SSVEP (Ross et al., 1989).
Others? Baker recent gain control paper is perhaps worth mentioning here
as a `confound' of sorts. Engel 2018
\href{https://www.zotero.org/google-docs/?kdURiS}{\emph{(Vergeer et al.,
2018)}}. This paper is interesting
\href{https://www.zotero.org/google-docs/?J2EO3d}{\emph{(Rideaux et al.,
2023)}} and \textgreater sort\textless{} of SSVEP.

\phantomsection\label{anchor-8}{}Masking:

High contrast masks can suppress the visibility of low contrast
patterns, which has implications in SSVEP amplitude (Haynes et al.,
2003).

Attention: Directing attention can enhance contrast sensitivity, as
shown in studies using SSVEP (Müller et al., 2006). Also Tsai
(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
dynamic model of normalization with some MEG data. Busse et al cat/human
comparison. Candy and Norcia 2001 JNS
\href{https://www.zotero.org/google-docs/?7hhs5Z}{(Candy et al., 2001)}

Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
adaptation and masking on incremental thresholds for contrast. Vision
research, 29(2), 205-215.

Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
Neuromagnetic correlates of perceived contrast in primary visual cortex.
Journal of Neurophysiology, 89(6), 2655-2666.

Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
selective attention on the steady-state visual evoked potential in the
20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.

5. Clinical implications

\phantomsection\label{anchor-9}{}Clinical applications (this is a whole
section)

Not sure exactly how SSVEP used in clinic. mfVEP?

, e.g., in monitoring visual impairments, tracking neuronal diseases, or
neurofeedback (Norcia et al., 2015).

Tsai epilepsy \href{https://www.zotero.org/google-docs/?AdhH3f}{(Tsai et
al., 2011)}. Other photosensitive epilepsy: Porciatti
\href{https://www.zotero.org/google-docs/?UTBQbV}{(Porciatti et al.,
2000)}

Citation:

Migraine (Regan et al). Autism?

Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
Rossion, B. (2015). The steady-state visual evoked potential in vision
research: A review. Journal of vision, 15(6), 4-4.

Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
(probably many - can't think off the top of my head - we were working
with a mouse EEG person at UCSF about 12 years ago...), Monkeys
(Kiorpes?),

\phantomsection\label{anchor-10}{}6. Future directions

Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
amblyopia in Rats,

GABA Huang

There is a lot of SSVEP interest these days because of BCIs. I think
it's pretty weak but there is \textgreater so much\textless{} of it that
it might be worth mentioning...

Advanced signal processing techniques and machine learning can be
integrated to improve SSVEP-based systems (Zhu et al., 2010).

Exploring new clinical and diagnostic applications, understanding
neurological diseases, and developing novel therapeutic interventions.

Citation:

Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
stimulation methods used in SSVEP-based BCIs. Computational intelligence
and neuroscience, 2010.

\section{References}\label{references}

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

The complexity of even a simple simulation of the frequency-domain
signal arising from non-linear interactions is intriguing. Presumably,
the signal measured from early visual cortex is the result of a cascade
of nonlinear retinal and cortical operations up to that point. It
therefore contains a `signature' or `fingerprint' of the computational
nature, order and parameters of those operations - including the shape
of the transducer functions and the computations involved in signal
combination. In principle, that information could be recovered from the
SSVEP signal - a possibility recognised in the early days of the
technique (Regan \& Regan, 1988). Although characterising the complete
set of computations along entire processing pathway is challenging,
careful parametric variation of the input stimuli does allow us to fit
models of early visual gain control using SSVEP data. For example, Tsai
\emph{et al.} (2012) demonstrated that a gain control model gave a good
account of the pattern of intermodulation responses produced by two
overlaid patterns flickering at different frequencies. This was achieved
by passing full stimulus waveforms through the transducer nonlinearity,
and calculating the Fourier spectrum of the model output. Our own work
on signal combination across eyes and space similarly demonstrated close
correspondence between the predictions of a computational model and
empirical data in humans (Baker \& Wade, 2017). More detailed modelling
of intracortical recordings (Groen \emph{et al.}, 2022) has revealed
details of the timecourse of gain control effects, specifically that
normalization is delayed slightly relative to the initial visual
response.

@@

Because SSVEP responses at detection threshold are very small,
estimating a threshold is achieved by measuring the contrast response
function at relatively high levels, and extrapolaKting back along the
function (either contrast vs response measured at a constant spatial
frequency or spatial frequency vs response at a constant contrast level)
to estimate its intercept with the x-axis (see
Figure~\ref{fig-sweepvep}). This contrast level was shown to correspond
approximately with psychophysically measured detection thresholds
(Norcia \emph{et al.}, 1986).

@@

\section{Attention and arousal}\label{attention-and-arousal}

Attention exerts a profound influence on visual performance: for
example, allowing people to attend covertly to a spatial location
improves their performance on a target detection task significantly
(Bashinski \& Bacharach, 1980; Posner, 1980; Carrasco \emph{et al.},
2000; Cameron \emph{et al.}, 2002; Morrone \emph{et al.}, 2004; Pestilli
\emph{et al.}, 2009). In principle, this enhancement may be driven both
by modulation of the underlying signal or noise characteristics, or by
additional decision-theoretic factors such as reduction in spatial
uncertainty (Petrov \emph{et al.}, 2006; Gould \emph{et al.}, 2007).
Early experiments with non-human primates showed little evidence for
attentionally-driven changes in neuronal spike rates (Luck \emph{et
al.}, 1997; McAdams \& Maunsell, 1999; Mehta \emph{et al.},
2000\emph{a}, 2000\emph{b}; Marcus \& Van Essen, 2002), but with the
advent of spatially-resolved human brain imaging methods in the late
1990s it became apparent that spatial attention was linked to frank
changes in both fMRI BOLD responses (Tootell \emph{et al.}, 1998; Somers
\emph{et al.}, 1999; Gandhi \emph{et al.}, 1999; Brefczynski \& DeYoe,
1999; Kastner \emph{et al.}, 1999; Buracas \& Boynton, 2007; Silver
\emph{et al.}, 2007; Li \emph{et al.}, 2008; Murray, 2008) and EEG
signals (Morgan \emph{et al.}, 1996; Müller \emph{et al.}, 1998; Müller
\& Hillyard, 2000; Ding \emph{et al.}, 2006).

Electrophysiological and psychophysical measurements of the effect of
attention on both luminance and chromatic contrast have strongly
implicated gain control as an underlying mechanism (Lu \& Dosher, 1998;
Di Russo \& Spinelli, 1999\emph{a}; Di Russo \emph{et al.},
2001\emph{b}). These effects also appear to differ between chromatic and
achromatic pathways (Di Russo \& Spinelli, 1999\emph{b}) - perhaps as a
result of the different levels of nonlinear gain control in the early
pre-cortical magno-, parvo- and konio-cellular pathways (Derrington \&
Lennie, 1984; Kaplan \& Shapley, 1986; Lee \emph{et al.}, 1990; Solomon
\& Lennie, 2005). In the late 2000s a comprehensive theoretical model
for attentional modulation was developed that framed it as a gain
control computation (Reynolds \& Heeger, 2009; Boynton, 2009). This
framework has proven to be influential - explaining a wide range of
phenomena from the earlier literature and demonstrating subtle
interactions between the size of the attentional `spotlight' and the
stimulus configuration which rationalise many apparent contradictions in
the literature. Direct measurements of attentional modulation of
achromatic SSVEP signals are broadly consistent with this model
(Lauritzen \emph{et al.}, 2010; Hou \emph{et al.}, 2016; Martinovic \&
Andersen, 2018) and confirm the relatively weaker role of spatial
attention on responses driven by chromatic stimuli - particularly those
that isolate the opponent S-(L+M) cone pathway (Highsmith \& Crognale,
2010; Wang \& Wade, 2011). The gain control model of attention can also
be extended to SSVEP studies of feature-based attention which show that
the modulatory effects can be targeted to the most informative neuronal
populations (Verghese \emph{et al.}, 2012).

The SSVEP can also be used to study changes in visual processing
\textit{driven} by different behavioural states or overall arousal. For
example, locomotion has been shown to alter neuronal excitability and
spatial normalization in mice (Niell \& Stryker, 2010; Ayaz \emph{et
al.}, 2013) - running mice have higher visual sensitivity and lower
surround suppression compared to stationary mice. Although measuring EEG
responses from locomoting humans is technically challenging, SSVEP
studies (which are able to distinguish broadband noise from input signal
effectively) have shown that walking also alters early visual
processing, although in a manner different to that observed in mice
(Benjamin \emph{et al.}, 2018; Cao \& Händel, 2019).

SSVEP measurements are typically used to measure time-invariant
responses due to sustained attention. However, recent work has shown
that moderately high modulation frequencies (ca 10Hz) and short analysis
windows (1.5s) can also be used to track the dynamic allocation of
attention across a task (Chota \emph{et al.}, 2024) or attention to
moving targets (Lissa \emph{et al.}, 2020). The ability to track changes
in attention over short time periods is also important if SSVEP is to be
used for dynamic readouts - for example in a Brain Computer Interface.

@@

One widespread recent application of the SSVEP technique is in the
design of brain-computer interfaces, which seek to control some aspect
of a computer using neural signals. The high SNR and precise frequency
resolution of SSVEPs make them an ideal candidate for this approach.
Typical studies may involve presenting an array of stimuli at different
contrast flicker frequencies, and having the participant select one
either by overt attention (i.e.~shifting fixation to foveate the
selected stimulus) or covert attention (i.e.~deploying attention to one
stimulus whilst keeping fixated) (Middendorf \emph{et al.}, 2000).
Because SSVEP signals are highly sensitive to both visual field position
(Di Russo \emph{et al.}, 2007; Ales \emph{et al.}, 2010) and attentional
state (Morgan \emph{et al.}, 1996; Müller \emph{et al.}, 1998; Lauritzen
\emph{et al.}, 2010; Verghese \emph{et al.}, 2012), the response to the
selected stimulus will typically increase relative to the others,
allowing it to be identified by an on-line algorithm. Because more than
one stimulus frequency is generally present, this modulation and the
associated changes in gain control will affect the entire pattern of
self- and intermodulation terms, allowing the choice to be decoded by a
multivariate pattern classifier.

This approach is primarily useful in situations that require the BCI to
distinguish from among a small set of possibilties: for example, in
early work, visual stimuli representing `Left' and `Right' commands in a
flight simulator were distinguished robustly (Middendorf \emph{et al.},
2000). Although SSVEP-based BCI interfaces typically do use contrast
flicker, work in this field has largely focused on optimising the
stimuli or decoders to increase decoding performance (e.g.~\ldots{} )
and reducing visual fatigue associated with the long-term presentation
of arrays of high-contrast flicker (Diez \emph{et al.}, 2024) rather
than studying visual contrast processing \textit{per se}. We therefore
note in passing that using SSVEPs to improve our understanding of early
contrast processing may yield benefits to this related field.

\section{Future directions}\label{future-directions}

The SSVEP is a powerful tool for studying contrast processing. It
provides a high SNR readout of neuronal activity that is unambigously
linked to the input. It is sensitive to both the amplitude and phase of
the input and when combined with source imaging, it can be extracted
from different cortical regions allowing researchers to track contrast
processing computations across the visual pathway. Because of the high
SNR, it can be measured in subjects where long recording durations are
impractical (for example, infants or patients with neurological
disorders) and at frequencies high enough to be effectively invisible
(Herrmann, 2001; Seijdel \emph{et al.}, 2023; Minarik \emph{et al.},
2023). The SSVEP is also able to `fingerprint' modulators of the inputs
through the harmonic and intermodulation terms they generate in the
output. In principle, each nonlinearity in the visual pathway can be
identified by its contribution to the frequency spectrum at different
recording locations (Regan \& Regan, 1988). This, it turn, allows
researchers to study how contrast processing depends on spatial and
temporal context and as well as changes in task, cognitive and
behavioural state and arousal.

Although visual neuroscience is a relatively old subfield, there are
still outstanding questions relating to contrast processing that could
be addressed by SSVEP methods. First, it is still not completely clear
how contrast signals are computed in the human retina. Although we have
more than a century of electrophysiological data from animals, and the
broad structure of cone inputs to retinal ganglion cells is understood
(Li \emph{et al.}, 2014), we are still discovering new aspects of
retinal processing that could influence the `coining' of the visual
system's currency (Gollisch \& Meister, 2010; Uprety \emph{et al.},
2022; Wang \emph{et al.}, 2023). The ability to frequency tag both
inputs (for example, cone-directed luminance contrast) and modulators
(for example, stimuli that selectively drive the
intrinsically-photosensitive retinal ganglion cells) is a powerful tool
to explore this first stage of image generation.

Later in the visual pathway, we would like to know more about the role
of corticothalamic feedback in the LGN. Although it is commonly thought
of as a simple relay between the eye and the cortex, the majority of
inputs to the LGN come from cortex not the eye, and there is good
evidence that contrast processing in the LGN can be altered by top-down
signals including attention (Sherman \& Guillery, 1996; O'Connor
\emph{et al.}, 2002; Briggs \& Usrey, 2007; Gouws \emph{et al.}, 2014).
Bottom-up inputs to the LGN are segregated by eye and so responses there
are often considered to be purely monocular, but it is possible that
feedback signals allow some binocular computations such as interocular
normalisation (Baker \emph{et al.}, 2007; Dougherty \emph{et al.}, 2019)
to begin even at this relatively early stage. Frequency tagging inputs
to different eyes or in different precortical pathways allows us to
address neurons in different parts of the LGN. These techniques combined
with advances in recording technology (for example, sensitive
source-imaged EEG and MEG recordings that can resolve subcortical
structures (Tesche, 1996; Attal \emph{et al.}, 2012; Attal \& Schwartz,
2013), noninvasive deep-brain stimulation techniques (Mohammadjavadi
\emph{et al.}, 2022) or implanted electrode arrays (Krolak-Salmon
\emph{et al.}, 2003)) may allow us to study these computations in more
detail.

Finally, SSVEPs continue to provide a way of studying contrast in the
cortex. Here, we are often interested in how different visual parameters
interact. For example, how are signals from different eyes combined to
generate both scalar contrast values and also binocular depth cues? How
does contrast combination depend on the low level properties of the
individual inputs such as their retinotopic location, cone contrast,
spatiotemporal frequency, eye of origin and orientation? Colour vision
scientists are particularly interested in how chromatic signals
originating in a small number of cone-opponent retinal pathways are
transformed into a perceptual colour space where the `unique hues'
appear to be only weakly-related to the early retinal outputs and how
these computations are conditioned by the spatial and temporal
properties of the scene (Wandell, 1993; Gegenfurtner, 2003; Solomon \&
Lennie, 2007; Stoughton \& Conway, 2008; Kaneko \emph{et al.}, 2020; Li
\emph{et al.}, 2022). All of these questions can be addressed by using
SSVEP and frequency tagging to examine the computations that combine and
transform the inputs at different cortical stages (Busse \emph{et al.},
2009; Baker \& Wade, 2017; Katyal \emph{et al.}, 2018; Chen \&
Gegenfurtner, 2021; Watts \emph{et al.}, 2024). The original promise of
the SSVEP approach was that the entire complex-valued frequency spectrum
recorded at each lcoation provided detailed information about the
processing nonlinearities up to and including that point. In principle,
this would allow researchers to uniquely fit the parameters of their
computations models of visual processing (Regan \& Regan, 1988). To
date, the complexity of the neuronal computations at even the earliest
stage of visual processing have hampered this effort - researchers
typically restrict their analyses to the amplitudes of single, low-order
frequency components (for example, the inputs frequencies or simple sums
and differences of those frequencies). As our understanding of the early
visual system improves, it is becoming possible to generate more
realistic parameterised forward models of signal generation (Tsai
\emph{et al.}, 2012; Baker \& Wade, 2017; Chariker \emph{et al.}, 2020;
Schrimpf \emph{et al.}, 2020; Groen \emph{et al.}, 2022). Feeding
frequency-tagged inputs into these types of model allows us to generate
synthetic SSVEP responses that can be compared with those measured in
human subject. In principle, we are therefore able to use the SSVEP to
fit the parameters of early visual processing. This approach may also
allow us to develop more sensitive tests for the changes in early visual
processing that accompany a wide range of neurological diseases and
disorders.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

@@

\bibitem[\citeproctext]{ref-Attal2012}
Attal Y, Maess B, Friederici A \& David O (2012).
\href{https://doi.org/10.1515/rns.2011.056}{Head models and dynamic
causal modeling of subcortical activity using
magnetoencephalographic/electroencephalographic data}. \emph{Rev
Neurosci} \textbf{23,} 85--95.

\bibitem[\citeproctext]{ref-Attal2013}
Attal Y \& Schwartz D (2013).
\href{https://doi.org/10.1371/journal.pone.0059856}{Assessment of
subcortical source localization using deep brain activity imaging model
with minimum norm operators: A MEG study}. \emph{PLoS One} \textbf{8,}
e59856.

\bibitem[\citeproctext]{ref-Ayaz2013}
Ayaz A, Saleem AB, Schölvinck ML \& Carandini M (2013).
\href{https://doi.org/10.1016/j.cub.2013.04.012}{Locomotion controls
spatial integration in mouse visual cortex}. \emph{Curr Biol}
\textbf{23,} 890--894.

@@

\bibitem[\citeproctext]{ref-Benjamin2018}
Benjamin AV, Wailes-Newson K, Ma-Wyatt A, Baker DH \& Wade AR (2018).
\href{https://doi.org/10.1523/JNEUROSCI.1428-17.2017}{The {Effect} of
{Locomotion} on {Early} {Visual} {Contrast} {Processing} in {Humans}}.
\emph{Journal of Neuroscience} \textbf{38,} 3050--3059.

@@

\bibitem[\citeproctext]{ref-Briggs2007}
Briggs F \& Usrey WM (2007).
\href{https://doi.org/10.1523/JNEUROSCI.1035-07.2007}{A fast, reciprocal
pathway between the lateral geniculate nucleus and visual cortex in the
macaque monkey}. \emph{J Neurosci} \textbf{27,} 5431--5436.

@@

\bibitem[\citeproctext]{ref-Cao2019}
Cao L \& Händel B (2019).
\href{https://doi.org/10.1371/journal.pbio.3000511}{Walking enhances
peripheral visual processing in humans}. \emph{PLoS Biol} \textbf{17,}
e3000511.

@@

\bibitem[\citeproctext]{ref-Chariker2020}
Chariker L, Shapley R \& Young L-S (2020).
\href{https://doi.org/10.1167/jov.20.4.16}{Contrast response in a
comprehensive network model of macaque V1}. \emph{J Vis} \textbf{20,}
16.

@@

\bibitem[\citeproctext]{ref-Chen2021}
Chen J \& Gegenfurtner KR (2021).
\href{https://doi.org/10.1167/jov.21.8.12}{Electrophysiological evidence
for higher-level chromatic mechanisms in humans}. \emph{J Vis}
\textbf{21,} 12.

\bibitem[\citeproctext]{ref-Chota2024}
Chota S, Bruat AT, Van der Stigchel S \& Strauch C (2024).
\href{https://doi.org/10.1162/jocn_a_02107}{Steady-state visual evoked
potentials reveal dynamic (re)allocation of spatial attention during
maintenance and utilization of visual working memory}. \emph{J Cogn
Neurosci} \textbf{36,} 800--814.

@@

\bibitem[\citeproctext]{ref-Diez2024}
Diez P, Orosco L, Garcés Correa A \& Carmona L (2024).
\href{https://doi.org/10.1007/s11517-023-03000-z}{Assessment of visual
fatigue in SSVEP-based brain-computer interface: A comprehensive study}.
\emph{Med Biol Eng Comput} \textbf{62,} 1475--1490.

@@

\bibitem[\citeproctext]{ref-Dougherty2019}
Dougherty K, Schmid MC \& Maier A (2019).
\href{https://doi.org/10.1002/cne.24417}{Binocular response modulation
in the lateral geniculate nucleus}. \emph{J Comp Neurol} \textbf{527,}
522--534.

@@

\bibitem[\citeproctext]{ref-Gegenfurtner2003}
Gegenfurtner KR (2003). \href{https://doi.org/10.1038/nrn1138}{Cortical
mechanisms of colour vision}. \emph{Nat Rev Neurosci} \textbf{4,}
563--572.

\bibitem[\citeproctext]{ref-Gollisch2010}
Gollisch T \& Meister M (2010).
\href{https://doi.org/10.1016/j.neuron.2009.12.009}{Eye smarter than
scientists believed: Neural computations in circuits of the retina}.
\emph{Neuron} \textbf{65,} 150--164.

@@

\bibitem[\citeproctext]{ref-Gouws2014}
Gouws AD, Alvarez I, Watson DM, Uesaki M, Rodgers J \& Morland AB
(2014). \href{https://doi.org/10.1523/JNEUROSCI.0164-14.2014}{On the
role of suppression in spatial attention: Evidence from negative BOLD in
human subcortical and cortical structures}. \emph{J Neurosci}
\textbf{34,} 10347--10360.

@@

\bibitem[\citeproctext]{ref-Herrmann2001}
Herrmann CS (2001). \href{https://doi.org/10.1007/s002210100682}{Human
EEG responses to 1-100 hz flicker: Resonance phenomena in visual cortex
and their potential correlation to cognitive phenomena}. \emph{Exp Brain
Res} \textbf{137,} 346--353.

@@

\bibitem[\citeproctext]{ref-Kaneko2020}
Kaneko S, Kuriki I \& Andersen SK (2020).
\href{https://doi.org/10.1093/texcom/tgaa059}{Steady-state visual evoked
potentials elicited from early visual cortex reflect both perceptual
color space and cone-opponent mechanisms}. \emph{Cereb Cortex Commun}
\textbf{1,} tgaa059.

@@

\bibitem[\citeproctext]{ref-Katyal2018}
Katyal S, Vergeer M, He S, He B \& Engel SA (2018).
\href{https://doi.org/10.1038/s41598-018-19809-w}{Conflict-sensitive
neurons gate interocular suppression in human visual cortex}. \emph{Sci
Rep} \textbf{8,} 1239.

\bibitem[\citeproctext]{ref-Krolak-Salmon2003}
Krolak-Salmon P, Hénaff M-A, Tallon-Baudry C, Yvert B, Guénot M,
Vighetto A, Mauguière F \& Bertrand O (2003).
\href{https://doi.org/10.1002/ana.10403}{Human lateral geniculate
nucleus and visual cortex respond to screen flicker}. \emph{Ann Neurol}
\textbf{53,} 73--80.

@@

\bibitem[\citeproctext]{ref-Li2022}
Li P, Garg AK, Zhang LA, Rashid MS \& Callaway EM (2022).
\href{https://doi.org/10.1038/s41467-022-34020-2}{Cone opponent
functional domains in primary visual cortex combine signals for color
appearance mechanisms}. \emph{Nat Commun} \textbf{13,} 6344.

\bibitem[\citeproctext]{ref-Li2014}
Li PH, Field GD, Greschner M, Ahn D, Gunning DE, Mathieson K, Sher A,
Litke AM \& Chichilnisky EJ (2014).
\href{https://doi.org/10.1016/j.neuron.2013.10.043}{Retinal
representation of the elementary visual signal}. \emph{Neuron}
\textbf{81,} 130--139.

@@

\bibitem[\citeproctext]{ref-Lissa2020}
Lissa P de, Caldara R, Nicholls V \& Miellet S (2020).
\href{https://doi.org/10.1371/journal.pone.0236967}{In pursuit of visual
attention: SSVEP frequency-tagging moving targets}. \emph{PLoS One}
\textbf{15,} e0236967.
